You Actually Look Twice At it (YALTAi): using an object detection approach instead of region segmentation within the Kraken engine

### 2022-07-19 article.tex
Preprint HAL

\abstract{Layout Analysis (the identification of zones and their classification) is the first step along line segmentation in Optical Character Recognition and similar tasks. The ability of identifying main body of text from marginal text or running titles makes the difference between extracting the work full text of a digitized book and noisy outputs. We show that most segmenters focus on pixel classification and that polygonization of this output has not been used as a target for the latest competition on historical document, despite being the focus in the early 2010s. We propose to shift, for efficiency, the task from a pixel classification-based polygonization to an object detection using isothetic rectangles. We compare the output of Kraken and YOLOv5 in terms of segmentation and show that the later severely outperforms the first on small datasets (1110 samples and below). We release two datasets for training and evaluation on historical documents as well as a new package, YALTAi, which injects YOLOv5 in the segmentation pipeline of Kraken 4.1.}

@@

In recent years, automatic text extraction has become an important activity in digital philology and, in general, in corpus creation for historical documents. Few different engines share the software market: Transkribus and eScriptorium and its Kraken engine are probably the two best known platforms. As model recognition improves over the year, the question of quasi-automatic corpus extraction from digitized manuscripts and early printed material has become more of a question of appropriate tooling rather than HTR performances for specific fonts or languages and periods. On the side of Kraken, which has the advantage of being fully open-source, unlike Transkribus, the layout segmentation has suffered from poor performances perception by its users, specifically with small dataset (around a thousand sample for train/dev/test) or very small dataset (less than 300 for train and dev).

This results perception has slowed the improvement of automatic information extraction, where the main body of text could be drawn from a printed book or a manuscript by ignoring marginal texts such as footnotes or running title. Extracting main body of works could result in creating data minable corpora of unprecedented sizes. We know of three research projects whose goals are hindered by the inability of Kraken to be trained with small dataset:

@@

Historical document layout analysis has seen its first competition created in 2011 as a joint venture from the \textit{International Conference on Document Analysis and Recognition} (ICDAR2011) and the \textit{International Workshop on Historical Document Imaging and Processing} (HIP2011). The 2011 event limited itself to print documents and tackled both the region segmentation as well as the region classification operations. In the context of ICDAR2011's Historical Document Layout Analysis competition (HDLA2011) and in general, layout segmentation is understood as the operation of recognizing blocks of content from the background in a digitized document. Region classification, on the other hand, is understood as the qualification of the blocks found during layout segmentation into various different categories. During ICDAR2011, document for the competition spanned from the 17\textsuperscript{th} to the early 20\textsuperscript{th} century and were segmented using polygons, which is the norm for this competition.

@@

While POD2017 sets the objective as a bounding box detection task, both HBA and HisDoc-Layout-Comp treats the task as a pixel labelling one, a pixel can be assigned to multiple classes. Since 2017, deep learning models have emerged in the competition, specifically in HBA and HisDoc-Layout-Comp where CNN models shined. However, the shift from polygon detection at ICDAR2011 to pixel-based categorization at ICDAR 2019 creates a bias for the downstream task such as main body extraction: pixel categorization ignores the necessary serialization and clustering of said pixels into various zones that can then be filled with lines and content. This approach is specifically damageful in tools such as Kraken (v4.1.2) as it led to the creation of hacks, wich zones being qualified with \textit{ColumnEven} and \textit{ColumnOdd} systems in order to ensure the differentiation of columns in the clustering step. Note that even stronger models such as ARU-NET and implemented by \citet{gruning2018arunet} only computes pixel classification and does not polygonize the output in its official implementation\footnote{\url{https://github.com/TobiasGruening/ARU-Net}}.

@@

The experimental setup is focused on evaluating the gains in terms of accuracy of zone detection or segmentation in historical documents with ``complex'' layout. We compare YoloV5 models to Kraken segmentation model using the same datasets (described below) with fix split and using Mean Average Precision (mAP) as well as Average Precision (AP) for each class of categories that are relevant to the qualitative analysis of results.

@@

\begin{table}[ht]
\resizebox{\linewidth}{!}{%}
\begin{tabular}{l|rrrrrrrr}
\hline
       & Dataset  & Batch Size & Architecture & Size (MiB)      & GPU Power (Max W) & GPU RAM \% (Max)   & Training Time   & Prediction Time \\ \hline
Kraken & Table    & 1          & 640          & \textbf{5}      & 202.1             & \textbf{25.44 \%}  &  00:50          & -               \\
YoloV5 & Table    & 2          & 5x           & 175             & \textbf{157.2  }  & 67.85              &  \textbf{00:25} & -               \\\hline
Kraken & Segmonto & 1          & 1200         &         5       & 236.1             & 66.02 \%           &  06:27          & -               \\
YoloV5 & Segmonto & 2          & 5x           & 175             & 160.43            & 39.31 \%           &  \textbf{03:04} & 0.025s          \\
YoloV5 & Segmonto & 2          & 5n           & \textbf{3.9}    & \textbf{54.05  }  & \textbf{11.5 \%}   &  03:11          & \textbf{0.004s} \\ \hline
\end{tabular}%\
}
\caption{Other metrics related to Kraken and YoloV5. YoloV5x produces the heaviest model (175mb, 35 times the weight of the Kraken model) but uses at most 77\% of the maximum power draw of Kraken. Training time is generally doubled under Kraken but corresponds roughly to the batch size division. Prediction time is per image. For the Segmonto model, the maximum GPU RAM usage is halved for YoloV5N, while using a batch of 2, but is nearly three times the footprint of Kraken for the Table dataset. Training occured on a NVIDIA RTX2080TI with a rated 250W of powers and 11,018MiB of RAM on drivers 440.118.02 and CUDA 10.2.}
\label{tab:scores:configs}
\end{table}

>>>>

\abstract{Layout Analysis (the identification of zones and their classification) is the first step along line segmentation in Optical Character Recognition and similar tasks. The ability of identifying main body of text from marginal text or running titles makes the difference between extracting the work full text of a digitized book and noisy outputs. We show that most segmenters focus on pixel classification and that polygonization of this output has not been used as a target for the latest competition on historical document (ICDAR 2017 and onwards), despite being the focus in the early 2010s. We propose to shift, for efficiency, the task from a pixel classification-based polygonization to an object detection using isothetic rectangles. We compare the output of Kraken and YOLOv5 in terms of segmentation and show that the later severely outperforms the first on small datasets (1110 samples and below). We release two datasets for training and evaluation on historical documents as well as a new package, YALTAi, which injects YOLOv5 in the segmentation pipeline of Kraken 4.1.}

@@

In recent years, automatic text extraction has become an important activity in digital philology and, in general, in corpus creation for historical documents. Few different tools share the software market: Transkribus and eScriptorium (with its underlying and autonomous Kraken engine) are probably the two best known platforms. As model recognition improves over the year, the question of quasi-automatic corpus extraction from digitized manuscripts and early printed material has become more of a question of appropriate tooling rather than HTR performances for specific fonts or languages and periods. On the side of Kraken, which has the advantage of being fully open-source, unlike Transkribus, the layout segmentation \citep{kiessling2019badam} has suffered from poor performances user-perception, specifically with small dataset (around a thousand sample for train/dev/test) or very small dataset (less than 300 for train and dev).

This results perception has slowed the improvement of automatic information extraction, where the main body of text could be drawn from a printed book or a manuscript by ignoring marginal texts such as footnotes or running title. If the segmenter is not able to recognize columns (or separate them) in medieval manuscripts or does not distinguish main bodies of text from foot notes, the co-occurrences of the extracted text are fooled by a wrong interpretation of the layout. Extracting main body of works could result in creating data minable corpora of unprecedented sizes. We know of three research projects whose goals are hindered by the inability of Kraken to be trained with small dataset:

@@

Layout Analysis has been one of the most frequent task in competition in conferences such as \textit{International Conference on Document Analysis and Recognition} (ICDAR) and International Workshop on Historical Document Imaging and Processing (HIP). However, historical document layout analysis competition saw the light of the day in 2011, as a joint venture from ICDAR2011 and HIP2011. The 2011 event limited itself to print documents and focused both the region segmentation as well as the region classification operations. In the context of ICDAR2011's Historical Document Layout Analysis competition (HDLA2011) and in general, layout segmentation is understood as the operation of recognizing blocks of content from the background in a digitized document. Region classification, on the other hand, is understood as the qualification of the blocks found during layout segmentation into various different categories. During ICDAR2011, document for the competition spanned from the 17\textsuperscript{th} to the early 20\textsuperscript{th} century and were segmented using polygons, which is the norm for this competition.

@@

While POD2017 sets the objective as a bounding box detection task, both HBA and HisDoc-Layout-Comp treats the task as a pixel labelling one, a pixel can be assigned to multiple classes. Since 2017, deep learning models have emerged in the competition, specifically in HBA and HisDoc-Layout-Comp where CNN models shined.

However, the shift from polygon detection at ICDAR2011 to pixel-based categorization at ICDAR 2019 creates a bias for the downstream task such as main body extraction: pixel categorization ignores the necessary serialization and clustering of said pixels into various zones that can then be filled with lines and content. This approach is specifically damageful in tools such as Kraken (v4.1.2) as it led to the creation of ``hacks'', wich zones being qualified with \textit{ColumnEven} and \textit{ColumnOdd} systems in order to ensure the differentiation of columns in the clustering step\footnote{While undocumented, this was the approach taken by many project including LECTAUREP under the advises of the SCRIPTA team who is working on Kraken and eScriptorium.}. Note that even stronger models such as ARU-NET \citep{gruning2018arunet} only computes pixel classification and does not polygonize the output in its official implementation\footnote{\url{https://github.com/TobiasGruening/ARU-Net}}.

@@

The experimental setup is focused on evaluating the gains in terms of accuracy of zone detection or segmentation in historical documents with ``complex'' layout. We compare YoloV5 models to Kraken segmentation model using the same datasets (described below) with fix splits\footnote{Splits are available in the dataset published with the present article.} and using Mean Average Precision (mAP) as well as Average Precision (AP) for each class of categories that are relevant to the qualitative analysis of results.

@@

\begin{table}[ht]
\resizebox{\linewidth}{!}{%}
\begin{tabular}{l|rrrrrrrr}
\hline
       & Dataset  & Batch Size & Architecture & Size (MiB)      & GPU Power (Max W) & GPU RAM \% (Max)   & Training Time   & Prediction Time \\ \hline
Kraken & Table    & 1          & 640          & \textbf{5}      & 202.1             & \textbf{25.44 \%}  &  00:50          & -               \\
YoloV5 & Table    & 2          & 5x           & 175             & \textbf{157.2  }  & 67.85              &  \textbf{00:25} & -               \\\hline
Kraken & Segmonto & 1          & 1200         &         5       & 236.1             & 66.02 \%           &  06:27          & -               \\
YoloV5 & Segmonto & 2          & 5x           & 175             & 160.43            & 39.31 \%           &  \textbf{03:04} & 0.025s          \\
YoloV5 & Segmonto & 2          & 5n           & \textbf{3.9}    & \textbf{54.05  }  & \textbf{11.5 \%}   &  03:11          & \textbf{0.004s} \\ \hline
\end{tabular}%\
}
\caption{Other metrics related to Kraken and YoloV5. YoloV5x produces the heaviest model (175mb, 35 times the weight of the Kraken model) but uses at most 77\% of the maximum power draw of Kraken. Training time is generally doubled under Kraken but corresponds roughly to the batch size division. Prediction time is per image. For the Segmonto model, the maximum GPU RAM usage is halved for YoloV5N, while using a batch of 2, but is nearly three times the footprint of Kraken for the Table dataset. Training and prediction occured on a NVIDIA RTX2080TI with a rated 250W of powers and 11,018MiB of RAM on drivers 440.118.02 and CUDA 10.2.}
\label{tab:scores:configs}
\end{table}
